---
title: "R Bridge Week 2 Assignment"
author: "Gabriel Maldonado"
date: "7/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


R Bridge Course Week 2 Assignment
One of the challenges in working with data is wrangling. In this assignment we will use R to perform this
task.
Here is a list of data sets: http://vincentarelbundock.github.io/Rdatasets/ (click on the csv index for a
list)


### 1. Use the summary function to gain an overview of the data set. Then display the mean and median for at least two attributes

```{r}
# Mean function 
calculate_mean <- function(numbers) {
  s = sum(numbers)
  N = length(numbers)
  # Calculate the mean
  result = s / N
  return (result)
}

# Median function
calculate_median <-function(numbers) {
  N = length(numbers)
  sort(numbers)
  
  # Fine the median
  if (N %% 2 == 0) {
    # if N is even, convert to int and match position
    m1 = as.integer(N / 2) - 1
    m2 = as.integer((N / 2) + 1) - 1
    median = (numbers[m1] + numbers[m2]) / 2
  } else {
    # Convert to integer, match position
    m = as.integer((N + 1) / 2) - 1
    median = numbers[m]
  }
  return (round(median))
}

# Import transaction data into R

mydata = read.csv("https://raw.githubusercontent.com/maldonag/KVGMV9AW8U/master/CUNY-SPS/R%20Bridge/Week2/Homework2/Transact.csv")  # read csv file from Github

transact <- data.frame(mydata)

summary(transact)

# Mean and median for attributes t1 and t2

t1mean <- calculate_mean(transact$t1)
t1median <- calculate_median(transact$t1)

t2mean <- calculate_mean(transact$t2)
t2median <- calculate_median(transact$t2)

paste("t1 mean is ", t1mean,", and t2 median is ", t1median)

paste("t2 mean is ", t2mean,", and t2 median is ", t2median)
```

### 2. Create a new data frame with a subset of the columns and rows. Make sure to rename it.

```{r}

x <- transact$X
t1 <- transact$t1
t2 <- transact$t2
time <- transact$time

transaction <- data.frame(subset(transact, t1 >= 0 & time >= 100))

```


### 3. Create new column names for the new data frame.

```{r}
colnames(transaction) <- c("Row", "Transaction1", "Transaction2", "Time")

```


### 4. Use the summary function to create an overview of your new data frame. The print the mean
and median for the same two attributes. Please compare.

```{r}
summary(transaction)

t1mean <- calculate_mean(transaction$Transaction1)
t1median <- calculate_median(transaction$Transaction1)

t2mean <- calculate_mean(transaction$Transaction2)
t2median <- calculate_median(transaction$Transaction2)

paste("t1 mean is ", t1mean,", and t2 median is ", t1median)

paste("t2 mean is ", t2mean,", and t2 median is ", t2median)
```


### 5. For at least 3 values in a column please rename so that every value in that column is renamed.For example, suppose I have 20 values of the letter “e” in one column. Rename those values so
that all 20 would show as “excellent”.

```{r}

# set transaction1 value type to character type
transaction$Transaction1 <- as.character(transaction$Transaction1)

# we find all values in transaction1 equal to 0, and we index it for replacement
transIndex <- transaction$Transaction1 == "0"
# We replace all the occurances of the index value with NULL 
transaction$Transaction1[transIndex] <- "NULL"

```


### 6. Display enough rows to see examples of all of steps 1-5 above.

```{r}
# first 100 rows and all columns
transaction[0:100,]

```


### 7. BONUS – place the original .csv in a github file and have R read from the link. This will be a very
useful skill as you progress in your data science education and career.

```{r}
### Please see #1
```

